Hysteresis in human computation:\\ how one task affects another

### 2015-01-14 docs/publications/2014.09_SCIENCE/man.tex
incorporate some revisions

Given the increasing prevalence of microtask-based research methods, there
is a desire to understand factors affecting the reliability of HPU
outputs.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training and pre-screening of 
workers \cite{le2010ensuring,paolacci2010running}, and user-interface 
design \cite{Finnerty2013}.  
Researchers have also investigated the effects of \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.

@@

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we performed a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias we found was insignificant 
(see Fig.~\ref{fig:theta}).  This suggests 
that prior task exposure is a stronger priming modality than framing.

@@

To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, during
the test tasks within
a given experiment, the food-primed workers would use food-related words 
more often than their non-food-primed counterparts.  However, this was not 
generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks
(\# food related words were identified using the wordnet corpus, augmented
with additional words obtained by crawling a recipe website (see supplementary
information for details)) (see Fig.~\ref{fig:specificity}).  This finding
rules out a seemingly-simple mechanism in which workers emphasize
content that had been present in earlier tasks.  Seeing content in prior tasks
affects the probability that a worker refers to it in subsequent tasks, but 
does not necessarily \textit{increase} the probability.

To deepen our understanding, we investigated workers' lexical richness in 
reference to food, that is, the number of \textit{unique} food-related words
used.  Even if workers provide an abundance of food-related words, if the 
same words are often repeated, there will of course be a smaller 
diversity.  This could happen, for example, if workers use generic references 
to food.
Both \textit{intertask} experiments showed that food-exposed workers had 
greater lexical richness, in reference to food, than their non-food-primed 
counterparts (as much as 20\% more) (see Fig.~\ref{fig:specificity}).  
The increase in lexical richness in the case of 
\textit{intertask-food-culture} is quite interesting, because in that 
experiment, food-primed workers made fewer total references to food.  
We also observed enrichment of the food-related lexicon in the 
\textit{priming-food-objects}, and \textit{echo-food-objects} experiments, 
although to a lesser extent.

@@

Prior to this study, little thought appears to have been given to the 
ordering and bundling of microtasks.  But our findings show that these are 
powerful design factors.  A natural response might be simply to randomize 
task ordering, a practice that is commonly employed.  But the sheer extent of 
bias that we observed leads us to question 
whether that is the best approach.  Although the bias would be random in 
that case, it is still on the order of 30-50\%, which is a major contribution
to noise in the results.  It would be preferable to understand intertask
effects in greater depth so that they might \textit{controlled} rather than
simply randomized.

@@

Here we have shown that a large bias can be introduced into microtask 
responses, simply from the performance of earlier tasks. 
These effects are much stronger than framing.  
While our findings raise serious concerns about the 
current state of microtask design, they also reveal potential opportunities 
for more 
refined control over worker focus and acuity.  Any application that relies
on the repetitive use of human judgment is likely subject to the phenomena
we described here.  Exploring the concerns and opportunities posed by 
intertask effects are important directions for future work, 
in the effort to develop performant and reliable methods for human 
computation.
\bibliography{newbib}
\bibliographystyle{Science}

>>>>

Given the increasing prevalence of microtask-based methods, there
is a desire to understand factors affecting the reliability of microtask
platforms.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training and pre-screening of 
workers \cite{le2010ensuring,paolacci2010running}, and user-interface 
design \cite{Finnerty2013}.  
Researchers have also investigated the effects of \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.

@@

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we performed a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias we found was statistically 
insignificant (\td{insert p-value})
(see Fig.~\ref{fig:theta}).  This suggests 
that prior task exposure is a stronger priming modality than framing.

@@

To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, during
the test tasks within
a given experiment, the food-primed workers would use food-related words 
more often than their non-food-primed counterparts.  However, this was not 
generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks
(\# food related words were identified using the wordnet corpus, augmented
with additional words obtained by crawling a recipe website (see supplementary
information for details)) (see Fig.~\ref{fig:specificity}).  This finding
rules out a seemingly-simple idea that workers emphasize
content that has been present in earlier tasks: seeing content in prior tasks
affects the probability that a worker refers to it in subsequent tasks, but 
does not necessarily \textit{increase} the probability.

To deepen our understanding, we investigated workers' lexical richness in 
reference to food, that is, the number of \textit{unique} food-related words
used.  Even if workers provide an abundance of food-related words, if the 
same words are often repeated, there will be less diversity.  
This could happen, for example, if workers use generic references 
to food.
Both \textit{intertask} experiments showed that food-exposed workers had 
greater lexical richness, in reference to food, than their non-food-primed 
counterparts (as much as 20\% more) (see Fig.~\ref{fig:specificity}).  
The increase in lexical richness in the case of 
\textit{intertask-food-culture} is particularly noteworthy, because in that 
experiment, food-primed workers made fewer total references to food.  
We also observed enrichment of the food-related lexicon in the 
\textit{priming-food-objects}, and \textit{echo-food-objects} experiments, 
although to a lesser extent.

@@

Prior to this study, little thought appears to have been given to the 
ordering and bundling of microtasks.  But our findings show that these are 
powerful design factors.  A natural response might be simply to randomize 
task ordering, a practice that is commonly employed.  But the sheer extent of 
bias that we observed leads us to question 
whether that is the best approach.  Although the bias would be random in 
that case, our results suggest that it is on the order of 30-50\%, 
which is a major contribution to noise in the results.
It would be preferable to understand intertask
effects in greater depth so that they might \textit{controlled} rather than
simply randomized.

@@

Here we have shown that a large bias can be introduced into microtask 
responses, simply from the performance of earlier tasks. 
These effects are much stronger than framing.  
While our findings raise serious concerns about the 
current state of microtask design, they also reveal potential opportunities 
for more 
refined control over worker focus and acuity.  Any application that relies
on the repetitive use of human judgment is likely subject to the phenomena
we described here.  Exploring the issues and opportunities posed by 
intertask effects are important directions for future work, 
in the effort to develop performant and reliable methods for human 
computation.
\bibliography{newbib}
\bibliographystyle{Science}
